---
title: "iteration_and_listcols"
output: github_document
date: "2025-10-28"
---

```{r}
library(tidyverse)
library(rvest)
```

## Make a list

```{r}
#this has to be a list, can't be a dataframe, because each vector is of different length
l = 
  list(
    vec_numeric = 1:23,
    char_vec = c("Rachel"),
    mat = matrix(1:8, nrow = 2, ncol = 4),
    summary = summary(rnorm(1000, mean = 4))
  )

l

l[[1]] #gives just the first vector
l[["vec_numeric"]] #call it by name and get it returned
l$mat #another way of pulling stuff out, not the most preferred way

```

## Make a different list

```{r}

list_normals = 
  list(
    a = rnorm(30, mean = 3, sd = 1),
    b = rnorm(30, mean = 30, sd = 1),
    c = rnorm(30, mean = 3, sd = 10),
    d = rnorm(30, mean = -3, sd = 4)
  )

list_normals

```

(copy and paste the function from last time)

```{r}
mean_and_sd = function(x) {
  
  if(!is.numeric(x)) {
    error("The input x should be numeric")
  } 
  
  if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more values")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)
  
  tibble(
    mean = mean_x,
    sd = sd_x
  )
  
}
```

```{r}

mean_and_sd(list_normals[[1]])
mean_and_sd(list_normals[[2]])
mean_and_sd(list_normals[[3]])
mean_and_sd(list_normals[[4]])

```

But this is inefficient, maybe let's create a for loop so we don't have to separately type the code a bunch of time.

Use a loop to iterate!

```{r}

output = vector("list", length = 4)

for (i in 1:4) {
  
  output[[i]] = mean_and_sd(list_normals[[i]])
  
}

output
#same effect as code above but way more efficient
```

Use `map` to do the same thing.

```{r}
output = map(list_normals, mean_and_sd) 
#syntax: output = map(input, function)

#wow! even MORE efficient!

output = map(list_normals, median) #same thing, different function
```

Check out some `map` variants

```{r}
map_dfr(list_normals, mean_and_sd) #creates a dataframe out of the output

map_dfr(list_normals, mean_and_sd, .id = "sample") #keeps the id variable

map_dbl(list_normals, median) #if the output will be a single number can use map_dbl
```

##LIST COLUMNS

Try to put my list into a dataframe!!

```{r}
#can put these 2 things into a dataframe because they both have length 4
listcol_df =
  tibble(
    name = c("a", "b", "c", "d"),
    sample = list_normals
  )
```

Did this really work?

```{r}
pull(listcol_df, sample)

pull(listcol_df, name)
```

Can I apply `mean_and_sd`??

```{r}
mean_and_sd(pull(listcol_df, sample)[[1]])
mean_and_sd(pull(listcol_df, sample)[[2]])
mean_and_sd(pull(listcol_df, sample)[[3]])
mean_and_sd(pull(listcol_df, sample)[[4]])
```

Iterate using `map`

```{r}
map(pull(listcol_df, sample), mean_and_sd)
```

What if i want to add a column to my input dataframe, and that new column has the output?

Adding a column ...

```{r}
listcol_df =
  listcol_df |> 
  mutate(
    summary = map(sample, mean_and_sd)
  )

pull(listcol_df, summary)
```


USEFUL!! UNNEST

```{r}
listcol_df |> 
  select(-sample) |> 
  unnest(summary) #separates into 2 columns the mean and sd
```

## Revisit NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

nsduh_import = function(html, table_num) {
  
  data = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  data

}

nsduh_import(nsduh_html, table_num = 1)
nsduh_import(nsduh_html, table_num = 2)
nsduh_import(nsduh_html, table_num = 3)
```

Try this with a for loop:


```{r}
output = vector("list", length = 3)

for (i in 1:3) {
  
  output[[i]] = nsduh_import(nsduh_html, i)
  
}
```

Do this with `map`

```{r}
map(1:3, nsduh_import, html = nsduh_html) #have to add html argument
```












