---
title: "iteration_and_listcols"
output: github_document
date: "2025-10-28"
---

```{r}
library(tidyverse)
library(rvest)
```

## Make a list

```{r}
#this has to be a list, can't be a dataframe, because each vector is of different length
l = 
  list(
    vec_numeric = 1:23,
    char_vec = c("Rachel"),
    mat = matrix(1:8, nrow = 2, ncol = 4),
    summary = summary(rnorm(1000, mean = 4))
  )

l

l[[1]] #gives just the first vector
l[["vec_numeric"]] #call it by name and get it returned
l$mat #another way of pulling stuff out, not the most preferred way

```

## Make a different list

```{r}

list_normals = 
  list(
    a = rnorm(30, mean = 3, sd = 1),
    b = rnorm(30, mean = 30, sd = 1),
    c = rnorm(30, mean = 3, sd = 10),
    d = rnorm(30, mean = -3, sd = 4)
  )

list_normals

```

(copy and paste the function from last time)

```{r}
mean_and_sd = function(x) {
  
  if(!is.numeric(x)) {
    error("The input x should be numeric")
  } 
  
  if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more values")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)
  
  tibble(
    mean = mean_x,
    sd = sd_x
  )
  
}
```

```{r}

mean_and_sd(list_normals[[1]])
mean_and_sd(list_normals[[2]])
mean_and_sd(list_normals[[3]])
mean_and_sd(list_normals[[4]])

```

But this is inefficient, maybe let's create a for loop so we don't have to separately type the code a bunch of time.

Use a loop to iterate!

```{r}

output = vector("list", length = 4)

for (i in 1:4) {
  
  output[[i]] = mean_and_sd(list_normals[[i]])
  
}

output
#same effect as code above but way more efficient
```

Use `map` to do the same thing.

```{r}
output = map(list_normals, mean_and_sd) 
#syntax: output = map(input, function)

#wow! even MORE efficient!

output = map(list_normals, median) #same thing, different function
```

Check out some `map` variants

```{r}
map_dfr(list_normals, mean_and_sd) #creates a dataframe out of the output

map_dfr(list_normals, mean_and_sd, .id = "sample") #keeps the id variable

map_dbl(list_normals, median) #if the output will be a single number can use map_dbl
```

##LIST COLUMNS

Try to put my list into a dataframe!!

```{r}
#can put these 2 things into a dataframe because they both have length 4
listcol_df =
  tibble(
    name = c("a", "b", "c", "d"),
    sample = list_normals
  )
```

Did this really work?

```{r}
pull(listcol_df, sample)

pull(listcol_df, name)
```

Can I apply `mean_and_sd`??

```{r}
mean_and_sd(pull(listcol_df, sample)[[1]])
mean_and_sd(pull(listcol_df, sample)[[2]])
mean_and_sd(pull(listcol_df, sample)[[3]])
mean_and_sd(pull(listcol_df, sample)[[4]])
```

Iterate using `map`

```{r}
map(pull(listcol_df, sample), mean_and_sd)
```










