---
title: "simulation"
output: html_document
date: "2025-10-30"
---

```{r}
library(tidyverse)

source("sim_mean_sd.R")

set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

We can simulate by running our function

```{r}
sim_mean_sd(n_subj = 400)
```

Can I "verify" the central limit theorem?

```{r}
output = vector("list", length = 100)

for (i in 1:100) {
  
  output[[i]] = sim_mean_sd(30)
  
}

output |> 
  bind_rows() |>  #stacks up all the individual dataframes
  ggplot(aes(x = mu_hat)) +
  geom_density()
```

Try to repeat with a map statement

```{r}
sim_results_df = 
  expand_grid(
    sample_size = 30, #input list, use 30 every single time
    iter = 1:100 #do this 100 times
 ) |> 
  mutate(
    results = map(iter, sim_mean_sd, n_subj = 30) #run sim_mean_sd for every sample size in that column (in this case it's 30 every time, 100 times). saving the results as a dataframe for each of the 100 times you ran the sim.
  ) |> 
  unnest(results) #take all the df's and make it just the sample mean and sd
```

repeat with a bigger sample size

```{r}
sim_results_df = 
  expand_grid(
    sample_size = c(30, 60, 90, 120), #i'm doing 100 iterations for each sample size
    iter = 1:1000 
 ) |> 
  mutate(
    results = map(sample_size, sim_mean_sd)
  ) |> 
  unnest(results)
```

Let's look at this

```{r}
sim_results_df |> 
  mutate(
    sample_size = str_c("n = ", sample_size),
    sample_size = fct_inorder(sample_size)
  ) |> 
  ggplot(aes(x = sample_size, y = mu_hat)) +
  geom_violin()
```


Let's try to summarize

```{r}
sim_results_df |> 
  group_by(sample_size) |> 
  summarize(
    emp_mean = mean(mu_hat),
    emp_se = sd(mu_hat)
  )
```

## Simple linear regression

```{r}
sim_df = 
  tibble(
    x = rnorm(30, mean = 1, sd = 1),
    y = 2 + 3 * x + rnorm(30, 0, 1)
  )

sim_df |> 
  ggplot(aes(x = x, y = y)) +
  geom_point()

slr_fit = lm(y ~ x, data = sim_df)

coef(slr_fit)
```

turn this into a function

```{r}
sim_regression = function(n_subj, beta_0 = 2, beta_1 = 3) {
  
  sim_df = 
     tibble(
       x = rnorm(n_subj, mean = 1, sd = 1),
       y = beta_0 + beta_1 * x + rnorm(30, 0, 1)
     )

slr_fit = lm(y ~ x, data = sim_df)

tibble(
  beta0_hat = coef(slr_fit)[1],
  beta1_hat = coef(slr_fit)[2]
)
}

```

```{r}
sim_regression(n_subj = 30)
```

writing a for loop

```{r}
output = vector("list", length = 500)

for (i in 1:500) {
  
  output[[i]] = sim_regression(n_subj = 30) #n fixed at 30 each time
  
}

output |> 
  bind_rows()
```



```{r}
slr_sim_results_df =
  expand_grid(
    sample_size = 30,
    iter = 1:500
  ) |> 
  mutate(
    results = map(sample_size, sim_regression)
  ) |> 
  unnest(results)

slr_sim_results_df |> 
  ggplot(aes(x = beta0_hat, y = beta1_hat)) +
  geom_point()
```

## One more example!

The birthday problem - if you put a group of people in a room, what are the chances at least 2 people share a bday? How many people do you need to put in a room to get at least a 50% chance?

```{r}
birthdays = sample(1:365, 5, replace = TRUE) #there are 365 bdays, sample size 5, and then if you put a number once, you CAN pick it again. This is my data generating system

#is there a repeated bday yes or no?

repeated_bday = length(unique(birthdays)) < 5 # this is the analysis, checking if people have the same bday. If there are less unique bdays than people in the room, you know people share a bday

repeated_bday
```

put this in a function

```{r}
bday_sim = function (n_room) {
  
  birthdays = sample(1:365, n_room, replace = TRUE) 

  repeated_bday = length(unique(birthdays)) < n_room
  
  repeated_bday
  
}

bday_sim(20)
```

```{r}
bday_sim_results =
  expand_grid(
    bdays = 5:50,
    iter = 1:2500
  ) |> 
  mutate(
    result = map_lgl(bdays, bday_sim)
  ) |> 
  group_by(
    bdays
  ) |> 
  summarize(
    prob_repeat = mean(result)
  )
```

plot this

```{r}
bday_sim_results |> 
  ggplot(aes(x = bdays, y = prob_repeat)) +
  geom_point() +
  geom_line()
```











