---
title: "writing_functions"
output: github_document
date: "2025-10-23"
---

```{r}
library(tidyverse)
library(rvest)
library(readxl)
library(p8105.datasets)

set.seed(1)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Start small!

Everyone loves z scores.

```{r}
x_vec = rnorm(20, 10, sd = 3.5)

(x_vec - mean(x_vec)) / sd(x_vec)
```

Write a function to compute z scores.

```{r}
z_scores = function(x) {
  
  if(!is.numeric(x)) {
    error("The input x should be numeric")
  } #Tell R I want an error message if I try to input non-numeric argument
  
  if (length(x) < 5) {
    stop("Only compute z-scores when the input has 5 or more values")
  }
  
  z = (x - mean(x)) / sd(x)
  
  z #asking for output to be returned
  
}
```

Let's try our function ...

```{r}

z_scores(x = x_vec)

num_vec = rnorm(123, mean = 14, sd = 0.4)

z_scores(x = num_vec)

```

Let's break our function...

```{r, error = TRUE}
#error = TRUE tells R, yea there will be an error, deal with it and move on

z_scores(3) #--> gives a missing value. You need at least 2 numbers for computing SD

z_scores("my name is ren") #--> error message - can only take a mean of a numeric value
```

## Let's compute stuff

Let's compute and return the mean and sd of a numeric vector.

```{r}

mean_and_sd = function(x) {
  
  if(!is.numeric(x)) {
    error("The input x should be numeric")
  } 
  
  if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more values")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)
  
  tibble(
    mean = mean_x,
    sd = sd_x
  )
  
}

mean_and_sd(x_vec)

```

## Make up data

Let's *simulate* some data

```{r}

sim_df = 
  tibble(
    x = rnorm(n = 30, mean = 5, sd = 2)
  )

sim_df |> 
  summarize(
    mut_hat = mean(x),
    sigma_hat = sd(x)
  )

#each time I run this code I get a slightly different mean and sd because I keep generate different data in this simulation. But what if instead of hitting run on this a zillion times, I want a function that does it for me?
  
```

Write a function to do simulations.

The inputs are

* `n_subj` is number of subjects
* `mu` is the true mean
* `sigma` is the population sd

```{r}
sim_mean_sd = function(n_subj=30, mu = 3, sigma = 2){
  
  sim_df = 
   tibble(
     x = rnorm(n = n_subj, mean = mu, sd = sigma)
   )
 
 sim_df |> 
   summarize(
     mut_hat = mean(x),
     sigma_hat = sd(x)
    ) 
  
}
```

(we wrote this in a code chunk last time; now it's being sourced) 
now I can delete that code chunk above....

```{r}
source("sim_mean_sd.R")
```


Let's run this function

```{r}
sim_mean_sd(n_subj = 3800)

sim_mean_sd(mu = 48.4)
```

Import the LoTR data

```{r}

fellowship_ring = readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")

lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words") |> 
  mutate(race = str_to_lower(race)) |> 
  select(movie, everything()) 


```

Writing a function to import the datasets

```{r}
lotr_import = function(cell_range, movie_title) {
  
  df = read_excel("./data/LotR_Words.xlsx", range = cell_range) |>
  mutate(movie = movie_title)

}

fellowship = lotr_import(cell_range = "B3:D6", movie_title = "Fellowship")
two_towers = lotr_import(cell_range = "F3:H6", movie_title = "Two Towers")
return =     lotr_import(cell_range = "J3:L6", movie_title = "Return")

bind_rows(fellowship, two_towers, return)

```

Look at one more example

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj_year = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

data_marj_month = 
  nsduh_html |> 
  html_table() |> 
  nth(2) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

data_marj_first = 
  nsduh_html |> 
  html_table() |> 
  nth(3) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

Write an import function

```{r}

nsduh_import = function(html, table_num) {
  
  data = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  data

}

nsduh_import(nsduh_html, table_num = 1)
nsduh_import(nsduh_html, table_num = 2)

```








